Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-06-29 introduction.tex
introduce  \ifreview
  to toggle between reviewer comments/not, so as to make it easier to get a handle on
  net space used.
reduction in introduction.   WIP.

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 1: listing the key observations in one place, and in a
	consistent fashion. I found The 5 points listed in the first part of the
	rebuttal quite compelling, and strongly urge the authors to put them in
	the paper in a prominent location.
	\item REVIEWER 1: A glossary would be welcomed.
	\item REVIEWER 5: I would have preferred to see a more detailed list of
	contributions, including the novelty of the proposed approach as compared
	to prior work.
	\item I am ambivalent about the use of the term ``converging'' in the
	paper's title. While the paper does port workloads to high-performance
	computing resources, it does not deal with compute-oriented optimizations
	(parallelization), lowering communication, etc., which are often major
	concerns in HPC programming. In this sense, the paper is closer in spirit
	to how to port high-throughput apps to high-performance clusters. Agreed?
\end{enumerate}

@@

As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
(Run 4), it is anticipated that the data volumes that will need analyzing
will increase by factors of 10-100 compared to the current phase (Run 2).
Data will be larger in volume but will also require more sophisticated
computational processing. In spite of the impressive scale of the ATLAS
distributed computing system, demand for computing systems will significantly
outstrip current and projected supply.  There are multiple levels at which
this problem needs to be addressed: the utilization of emerging parallel
architectures (e.g., platforms); algorithmic and advances in analytical
methods (e.g., use of Machine Learning); and the ability to exploit different
platforms (e.g., clouds and supercomputers).

This paper represents the experience of how the ATLAS experiment has ``broken
free'' of the traditional computational approach of high-throughput computing
on distributed resources to embrace new platforms, in particular  high-
performance computers. Specifically, we discuss the experience of integrating
PanDA WMS with a US DOE leadership computing facility called Titan to reach
sustained production scales of approximately 51M core-hours a year. We also
discusses the investigation of a pilot-abstraction based task execution
runtime system on Titan, which allows advanced execution modes of the ATLAS
workload as well as the enhanced support for heterogeneous workloads, e.g.,
molecular dynamics.

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 1: listing the key observations in one place, and in a
	consistent fashion. I found The 5 points listed in the first part of the
	rebuttal quite compelling, and strongly urge the authors to put them in
	the paper in a prominent location.
	\item REVIEWER 1: A glossary would be welcomed.
	\item REVIEWER 5: I would have preferred to see a more detailed list of
	contributions, including the novelty of the proposed approach as compared
	to prior work.
	\item I am ambivalent about the use of the term ``converging'' in the
	paper's title. While the paper does port workloads to high-performance
	computing resources, it does not deal with compute-oriented optimizations
	(parallelization), lowering communication, etc., which are often major
	concerns in HPC programming. In this sense, the paper is closer in spirit
	to how to port high-throughput apps to high-performance clusters. Agreed?
\end{enumerate}
\fi

@@

In spite of the impressive scale of the ATLAS distributed computing system,
demand for computing systems will soon significantly outstrip current and
projected supply.   The data volumes that will need analyzing in LHC-Run 3 in
$~\approx$ 2022 and the high-luminosity era (Run 4) will increase by factors
of 10-100 compared to the current phase (Run 2). There are multiple levels at
which this problem needs to be addressed: the utilization of emerging parallel
architectures (e.g., platforms); algorithmic and advances in analytical
methods (e.g., use of Machine Learning); and the ability to exploit different
platforms (e.g., clouds and supercomputers).

% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity era
% (Run 4), it is anticipated that the data volumes that will need analyzing
% will increase by factors of 10-100 compared to the current phase (Run 2).
% Data will be larger in volume but will also require more sophisticated
% computational processing.

This paper represents the experience of how the ATLAS experiment has ``broken
free'' of the traditional computational approach of high-throughput computing
on distributed resources to embrace new platforms, in particular  
high-performance computers. Specifically, we discuss the experience of integrating
PanDA WMS with a US DOE leadership computing facility called Titan to reach
sustained production scales of approximately 51M core-hours a year. We also
discusses the investigation of a pilot-abstraction based task execution
runtime system on Titan, which allows advanced execution modes of the ATLAS
workload as well as the enhanced support for heterogeneous workloads, e.g.,
molecular dynamics.
